\MakeUppercase{The Structure of $k$-Planar Graphs} \thanks{This work was partly funded by NSERC and MRI.}

### 2019-07-05 k-planar.tex
Generalization is done

\begin{proof}
  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{k,d}$.  In particular, $G^{k,d}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.  To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.  This increases the thickness given by \clmref{general-width} from $O(k\ell d^k)$ to $O(k^2\ell d^k)$.
\end{proof}

It is natural to ask if \thmref{modern-major-general} is tight.  We do not answer this question, but we do point out that, even when $t=1$ and $d=3$, an exponential dependence on $k$ is necessary.  If $G$ is a complete binary tree on $2^{k/2}$ leaves, then $G^{k,3}$ is a clique of size  $s=2^{k/2+1}-1$.  Any partition of $G^{k,3}$ into $r$ parts will give an quotient graph $I$ that is an $r$-clique, which has treewidth $r-1$. On the other hand, any layering of $V(G^{k,3})$ has at most two layers, so the layered width of any partition into $r$ parts is at least $s/(2r)$. Therefore any layered $I$-partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
\[
    \ell'\times t' \ge 2^{k/2}-1/2 \enspace .
\]   
\newpage

>>>>

\begin{proof}
  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\mathcal{P}}$.  In particular, $G^{\mathcal{P}}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.  To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.  This increases the thickness given by \clmref{general-width} from $d\ell(k^2+k+4)/2$ to $d\ell(k^3+k^2+4k)/2$.
\end{proof}

% It is natural to ask if \thmref{modern-major-general} is tight.  We do not answer this question, but we do point out that, even when $t=1$ and $d=3$, an exponential dependence on $k$ is necessary.  If $G$ is a complete binary tree on $2^{k/2}$ leaves, then $G^{k,3}$ is a clique of size  $s=2^{k/2+1}-1$.  Any partition of $G^{k,3}$ into $r$ parts will give an quotient graph $I$ that is an $r$-clique, which has treewidth $r-1$. On the other hand, any layering of $V(G^{k,3})$ has at most two layers, so the layered width of any partition into $r$ parts is at least $s/(2r)$. Therefore any layered $I$-partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
\[
    \ell'\times t' \ge 2^{k/2}-1/2 \enspace .
\]   
\newpage
